---
title: "Compare equilibrium solution with Berkeley Madonna output"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If the equilibrium solution in this package is correct then it should be an exact match to the output of the Berkeley Madonna (BM) model run to equilibrium. Here we check this by importing BM model output for a range of EIR and treatment levels, calculating the equivalent quantities using the internal equilibrium solution, and comparing the two with some simple plots. 

Note that the BM output was produced using 164 age categories and 5 heterogeneity categories, and so the same values and breaks are used in the equilibrium solution.


## All model states

Here we explore the proportion of individuals in each model state as a function of EIR and treatment. First, read in BM output.

```{r, echo=FALSE}
# hard-coded to install package from Github or from file
rm(list=ls())
library(devtools)
if (FALSE) {
  install_github("mrc-ide/malariaModelFit", auth_token=my_token)  # use your own Github personal access token as my_token here. This step is required as the package is currently private.
  library(malariaModelFit)
} else {
  setwd('~/Dropbox/Bob/Work/My Programs/Personal packages/malariaModelFit')
  load_all('malariaModelFit')
}
```

```{r}
# define treatment levels corresponding to BM output files
ft <- c("0.0", "0.2", "0.4")

# read in BM output as list
BM_list <- list()
for (i in 1:length(ft)) {
  name <- paste0("BM_output/BM_ft",ft[i],".csv")
  BM_list[[i]] <- read.csv(malariaModelFit_file(name))
}

# get EIR range
EIR <- BM_list[[1]]$EIRY_eq
```

Now simulate the same output directly from the equilibrium solution.

```{r}
# load parameters
age <- default_age()
h <- gq_normal(5)
p <- load_parameters("parameters_Griffin2014.txt")

# save solution to list
eqList <- list()
for (i in 1:length(ft)) {
  eqList[[i]] <- list()
  for (j in 1:length(EIR)) {
      eqList[[i]][[j]] <- human_equilibrium_cpp(EIR[j], as.numeric(ft[i]), p, age, h$nodes, h$weights)
  }
}

# save state proportions in array
stateNames <- c("S", "T", "D", "A", "U", "P")
m <- array(0, dim=c(length(ft), length(EIR), length(stateNames)))
for (i in 1:length(eqList)) {
  for (j in 1:length(eqList[[i]])) {
    m[i,j,] <- colSums(subset(eqList[[i]][[j]]$states, select=stateNames))
  }
}

```

We can eyeball the two solutions by overlaying lines from the BM solution (solid) and from the internal equilibrium solution (dashed):

```{r, echo=FALSE, fig.width=8, fig.height=12}
# plot BM output and equilibrium solutions on same plot
par(mfrow=c(3,1))
for (i in 1:length(ft)) {
  BM_states <- as.list(BM_list[[i]])[-1]
  eq_states <- as.list(as.data.frame(m[i,,]))
  
  ns <- length(stateNames)
  s <- as.list(rep(c(1,2),times=rep(ns,2)))
  
  mplot(EIR, c(BM_states, eq_states), type="l", col=as.list(rep(1:ns,2)), lwd=s, lty=s, ymin=0, ymax=1, xlab="EIR", ylab="proportion", main=paste0("ft = ",ft[i]), legend_margin=8, legend1_title="Model state", legend1_label=stateNames, legend1_col=1:length(stateNames), legend1_pch=NULL, legend2_title="Simulation type", legend2_label=c("BM","equilibrium"), legend2_lty=c(1,2), legend2_lwd=c(1,2), legend2_pch=NULL, log="x")
}
```

Some proportions are too small to see cleary, so here are equivalent plots for the diseased state only:

```{r, echo=FALSE, fig.width=8, fig.height=10}
# plot BM output and equilibrium solutions on same plot
par(mfrow=c(3,1))
for (i in 1:length(ft)) {
  
  BM_states <- as.list(BM_list[[i]])$D
  eq_states <- as.data.frame(m[i,,])
  names(eq_states) <- stateNames
  eq_states <- eq_states$D
  
  mplot(EIR, list(BM_states, eq_states), type="l", col=2, lwd=list(1,2), lty=list(1,2), ymin=0, ymax=0.015, xlab="EIR", ylab="proportion", main=paste0("ft = ",ft[i]), legend_margin=8, legend1_title="Model state", legend1_label="D", legend1_col=2, legend1_pch=NULL, legend2_title="Simulation type", legend2_label=c("BM","equilibrium"), legend2_lty=c(1,2), legend2_lwd=c(1,2), legend2_pch=NULL, log="x")
  
}
```

(Rmarkdown last evaluated: `r Sys.time()`)